>Greenhouse Gas Emmision 
>by Ayush Dosajh


========================================================

> Resources -  

### http://datasets.wri.org/dataset/cait-country (for data set)

### www.datascience.com

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# install these if necessary
#install.packages('GGally')
#install.packages('scales')
#install.packages('memisc')
#install.packages('lattice')
#install.packages('MASS')
#install.packages('car')
#install.packages('reshape')
#install.packages('plyr')
#install.packages('RColorBrewer', dependencies = TRUE)


# load the ggplot graphics package and the others
library(RColorBrewer)
library(ggplot2)
library(GGally)
library(scales)
library(memisc)
library(gridExtra)
library(tseries)
library(forecast)
library(reshape2)
library(dplyr)




```

>Data overveiw/ summary

```{r echo=FALSE, message=FALSE, warning=FALSE ,  Load_the_Data}
# Load the Data


ghg <- read.csv("GHG.csv" , sep = ",")

summary(ghg)
#getwd()

```




> CAIT data set of global Green house gas emmision is acquired from - 
http://datasets.wri.org/dataset/cait-country .The CAIT Country GHG emissions collection applies a consistent methodology to create a six-gas, multi-sector, and internationally comparable data set for 186 countries.

>DATA names with its Units
 "Country" -Name of the countries   
 "Year"- year The data is for               
 "TotalGHG"- in Metric Tonnes         
 "Total.GHG.land"- in Metric Tonnes   
 "Total.CO2.land"- in Metric Tonnes   
 "TotalCH4"- in Metric Tonnes         
 "TotalN2O"- in Metric Tonnes          
 "TotalF.Gas"- in Metric Tonnes       
 "TotalCO2.land"- in Metric Tonnes    
 "TotalCH4.land"- in Metric Tonnes     
 "TotalN2O.land"- in Metric Tonnes     
 "Energy"- in Metric Tonnes           
 "Industrial"- in Metric Tonnes        
 "Agriculture"- in Metric Tonnes       
 "Waste"- in Metric Tonnes            
 "Land.Use" - in Metric Tonnes         
 "BunkerFuels"- in Metric Tonnes       
 "Electricity.Heat" - in Metric Tonnes
 "Manufacturing"- in Metric Tonnes     
 "Transportation"- in Metric Tonnes   
 "OtherFuel"- in Metric Tonnes        
 "FugitiveEmissions"- in Metric Tonnes



# Univariate Plots Section

> All the univariate plots are here.

A simple histogram dipicting TotalGHG distribution (it shows that most of the countries over the year has very less emmision also that there are more emission by certain coutries over the year other than this not much can be drawn out) 


```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots}

# A simple histogram dipicting TotalGHG distribution (it shows that most of the countries over the year has very less emmision also that there are more emission by certain coutries over the year other than this not much can be drawn out) 


qplot(x=TotalGHG ,data = ghg ) +
  scale_x_continuous(limits = c(0,1000) , breaks = seq(0,1000,100)) +
  scale_y_sqrt()


```

produces histograms of count(here no. countries with GHG Emmision for each year seperatly)

```{r echo=FALSE, message=FALSE, warning=FALSE , fig.width = 15}

#to know About distribution and trend over the years of Total GHG emmission (Total GHG is taken as one of the mai features)

qplot(x=TotalGHG,data = ghg ,ylab = 'NumberOfCountries', color = I('cyan'),
      fill= I('green')  )+
  scale_x_continuous(limits = c(100,1000) , breaks = seq(100,1000,100))+
  facet_wrap(~Year)

```

```{r echo=FALSE, message=FALSE, warning=FALSE , fig.width=15}

qplot(x=TotalCH4 , data = ghg ,ylab = 'NumberOfCountries'  )+
  scale_x_continuous(limits = c(100,1000) , breaks = seq(100,1000,100))+
  facet_wrap(~Year)

```

log trans to check for normal distribution

```{r echo=FALSE, message=FALSE, warning=FALSE}

#it is found out to be a bimodal normal distribution
qplot(x=TotalGHG ,data = ghg , color = I('cyan') , fill= I('red') ) +
  scale_x_continuous(limits = c(0,1000) , breaks = seq(0,1000,100))+
  scale_x_log10()

```



```{r echo=FALSE, message=FALSE, warning=FALSE , fig.width= 15}


qplot(x= Industrial ,data = ghg , color = I('cyan') , fill= I('red') ) +
  scale_x_continuous(limits = c(0,400) , breaks = seq(0,400,400))+
  scale_y_continuous(limits = c(0,20) , breaks = seq(0,50,10)) +
  facet_grid(~Year , space ='free_x' , scales = 'free_x')

```

 GHG PRODUCTION FROM DIFFRENT SOURCES and plotting side by side using grid arrange

```{r echo=FALSE, message=FALSE, warning=FALSE}

p1 <-qplot(x= Industrial ,data = ghg ,
           color = I('cyan') , fill= I('red') ) +
  scale_x_continuous(limits = c(0,400) , breaks = seq(0,500,200))+
  scale_y_continuous(limits = c(0,20) , breaks = seq(0,50,10)) 

p2 <-qplot(x= Transportation ,data = ghg , 
           color = I('cyan') , fill= I('red') ) +
  scale_x_continuous(limits = c(0,400) , breaks = seq(0,500,200))+
  scale_y_continuous(limits = c(0,20) , breaks = seq(0,50,10)) 

p3<- qplot(x= Waste ,data = ghg , 
           color = I('cyan') , fill= I('red') ) +
  scale_x_continuous(limits = c(0,250) , breaks = seq(0,500,200))+
  scale_y_continuous(limits = c(0,20) , breaks = seq(0,50,10)) 

p4<- qplot(x= Land.Use ,data = ghg , 
           color = I('cyan') , fill= I('red') ) +
  scale_x_continuous(limits = c(0,300) , breaks = seq(0,500,200))+
  scale_y_continuous(limits = c(0,20) , breaks = seq(0,50,10)) 

p5<- qplot(x= FugitiveEmissions ,data = ghg , 
           color = I('cyan') , fill= I('red') ) +
  scale_x_continuous(limits = c(0,400) , breaks = seq(0,500,200))+
  scale_y_continuous(limits = c(0,20) , breaks = seq(0,50,10)) 

grid.arrange(p1,p2,p3,p4,p5 , ncol=2)






```

> Univariate Findings

# Univariate Analysis


### What is the structure of your dataset?
The data set contains 4324 obs over 22 variables . Dataset gives info of major green house gasses and thier sources . 

### What is/are the main feature(s) of interest in your dataset?

The sources as well as the the total emmisions are given and hence contribution of each sector on total emmision.

### What other features in the dataset do you think will help support your \investigation into your feature(s) of interest?

Distribution of data over the years . or groping them by countries / continents to compare distributions.  

### Did you create any new variables from existing variables in the dataset?
No 

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?
No changes  were done to the data as it was clean and ready to use


# Bivariate Plots Section


```{r echo=FALSE , message=FALSE, warning=FALSE , Bivariate_Plots}

color.sub<- c( 'China' , 'Japan' , 'Bangladesh' , 'Pakistan' , 'India' , 'Afghanistan')
color.sub2<- c( 'China' , 'India' )
ghg1 <- subset (ghg, (ghg$Country == 'India' | ghg$Country == 'China' 
                      | ghg$Country == 'Japan'))


ggplot(aes(x=Total.CO2.land , y=TotalGHG ), data = ghg) +
  geom_line()

```

 ggpairs of selected country data only to display scatterplot matrix

```{r echo=FALSE, message=FALSE, warning=FALSE , fig.height= 12 , fig.width=15}

suppressWarnings(ggpairs(ghg1 , cardinality_threshold = 188  , upper = 'blank'))






```


 Almost linear relation ship between land + CH4 emmision and CH4 emmision but still there are some instances where CH4+ land is more than CH4

```{r echo = FALSE ,message=FALSE, warning=FALSE}

ggplot(aes(x=TotalCH4 , TotalCH4.land  ) , data = ghg) + 
  geom_point()+ 
  scale_x_continuous(breaks = seq(0,200,100) , limits = c(0,200)) +
  scale_y_continuous(breaks = seq(0,200,100) , limits = c(0,200))+
  stat_smooth(method = "lm")



```

 For most part japan's Emmision due to  energy production is 1000 which is far lower to China which Goes as far as about 9,000 Metric tonnes !!!!

```{r echo = FALSE , message=FALSE, warning=FALSE}
qplot(x= Energy , y = ..count.. / sum(..count..) , data = ghg1 ,
      xlab = 'Energy Emission' , ylab = 'count density'  , geom = 'freqpoly' , 
      color = Country ) +
  scale_x_continuous( breaks = seq(0,1000,10)) +
  scale_x_log10()

summary(ghg1)

```


Relationship between land + CO2 emmision and CO2 emmision is more varied though a bit of linear trend could be seen but it is not as linear as CH4 or N2O but rather it more spread around mean

```{r echo = FALSE ,message=FALSE, warning=FALSE}

ggplot(aes(x=Total.CO2.land , y = TotalCO2.land  ) , data = ghg) + 
  geom_point()+ 
  scale_x_continuous(breaks = seq(0,400,100) , limits = c(0,400)) +
  scale_y_continuous(breaks = seq(0,300,100) , limits = c(0,300))+
  stat_smooth(method = "lm")

ggplot(aes(x=TotalN2O , y = TotalN2O.land ) , data = ghg) + 
  geom_point()+ 
  scale_x_continuous(breaks = seq(0,400,100) , limits = c(0,400)) +
  scale_y_continuous(breaks = seq(0,300,100) , limits = c(0,300))+
  stat_smooth(method = "lm")






#names(ghg)

```


# Bivariate Analysis

>  Summarization - The variable are highly co-related and shows linear trend which is to be expected as all the date is about ghgs hence Year vs variable plots are taken up for analysis to produce meaningful plots


### Talk about some of the relationships you observed in this part of the \investigation. How did the feature(s) of interest vary with other features in \the dataset?
The features like Industrial emmision , Agriculture Emmision tend to increase over the years not much seasonality is seen the trend of emission of most variables are on the increasing side along the years.



### Did you observe any interesting relationships between the other features \(not the main feature(s) of interest)?

Relationship between land + CO2 emmision and CO2 emmision is more varied though a bit of linear trend could be seen but it is not as linear as CH4 Also N20 too doesnot show much daviation this could be either because Co2 has varied sources that effects its emmision and though land does effect Co2 but the effect is not purely linear .

### What was the strongest relationship you found?

Most of  relation ships are strong as the whole data set has data of Green house gasses


# Multivariate Plots Section

```{r echo=FALSE , message=FALSE, warning=FALSE}
ghg2 <-ghg1
ggplot(aes(x = Country, y = TotalGHG),data = ghg1) +
  geom_boxplot() +
  stat_summary( fun.y = mean , geom = 'point' , shape =8)



ghga <- subset(ghg ,   ghg$Country == 'India' |
                       ghg$Country == 'China' | ghg$Country == 'Japan')
ghga$year_bucket <- cut (ghga$Year , c(1990 ,1995 , 2000,2005 ,2012 ))
table(ghga$year_bucket, useNA = 'ifany')


ghga <- na.omit(ghga)

```

The points disply huge disparity b/w some of the observation  but the genral trend is exponential increse of industrial emminsion in comparison with Agricultral emminsion over the years.

```{r echo=FALSE , message=FALSE, warning=FALSE}


ggplot(aes(x= Agriculture , y= Industrial) ,
       data = ghga) +
  geom_line(aes(color = year_bucket) )

ggplot(aes(x= TotalGHG , y= Country ) ,
       data = ghga)+
  geom_line(aes(color = year_bucket) )



```

Emmision produced bu countries the trend is increasing  also ,Japan Emmsions are very punitive in comparison to China or india .the Graph is diffucult to understand as is and hence is updated to show the results clearly in "Final plot 1". 


PLOTTING ALL THE GASSES COUNTRY WISE USING GRIDEXTRA.gridarrange

```{r echo=FALSE , message=FALSE, warning=FALSE}



px <- ggplot(aes(x=Year , y= TotalGHG ,color = Country ),
             data = ghg1) +
  geom_line()+ ylab('totalGHG')



py <- ggplot(aes(x=Year , y= Total.CO2.land ,
                 color = Country ), data = ghg1) +
  geom_line()+
  ylab('CO2')

pz <- ggplot(aes(x=Year , y= TotalCH4 ,color = Country ),
             data = ghg1) +geom_line()+ ylab('CH4')

pa <- ggplot(aes(x=Year , y= TotalN2O ,color = Country ), 
             data = ghg1) +geom_line()+ ylab('N20')

grid.arrange(px,py,pz,pa ,ncol=1)






```

 PLOTTING ALL THE indutry  emmision COUNTRY WISE USING GRIDEXTRA.gridarrange

```{r echo=FALSE ,message=FALSE, warning=FALSE}

#PLOTTING ALL THE indutry  emmision COUNTRY WISE USING GRIDEXTRA.gridarrange
pl1 <- ggplot(aes(x=Year , y= Industrial ,color = Country ),
             data = ghg1) +
  geom_line()+ ylab('Industrial')



pl2 <- ggplot(aes(x=Year , y= Waste,
                 color = Country ), data = ghg1) +
  geom_line()+
  ylab('waste')

pl3 <- ggplot(aes(x=Year , y= Agriculture ,color = Country ),
             data = ghg1) +geom_line()+ ylab('Agri')

pl4 <- ggplot(aes(x=Year , y= Energy ,color = Country ), 
             data = ghg1) +geom_line()+ ylab('ENERGY')



pl5 <- ggplot(aes(x=Year , y= Manufacturing ,color = Country ), 
             data = ghg1) +geom_line()+ ylab('Manufacturing')


pl6 <- ggplot(aes(x=Year , y=Transportation  ,color = Country ), 
             data = ghg1) +geom_line()+ ylab('Transport')
lay <- rbind(c(1,1,1,2,2),
             c(1,1,1,2,2),
             c(1,1,1,2,2),
             c(3,3,3,4,4),
             c(3,3,3,4,4),
             c(5,5,5,6,6),
             c(5,5,5,6,6)
             )

grid.arrange(pl1,pl2,pl3,pl4,pl5,pl6 , ncol=1 , layout_matrix = lay)
```


 Heat map


```{r echo=FALSE, message=FALSE, warning=FALSE , Multivariate_Plots}
#heat map

ghg_heat_map <-subset(ghg , select = c( "Energy" , "Industrial" ,
                                        "Agriculture","Waste" ,"Land.Use" ,
                                        "Electricity.Heat" ,"Manufacturing" ,
                                        "Transportation" , "FugitiveEmissions"))
colnames(ghg_heat_map) <- c(1:9)

summary(ghg_heat_map)
dim(ghg_heat_map)
View(ghg_heat_map)
Long.2 <- melt(as.matrix(ghg_heat_map[24:92,]))

names(Long.2) <- c("countryYP", "v2", "value")

#HEAT MAP
ggplot(aes(y = countryYP, x = v2, fill = value),
  data = Long.2) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))





```



# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \investigation. Were there features that strengthened each other in terms of \looking at your feature(s) of interest?
China is mostly at the top for most variables / Analysis for emmision of variables . The trend is in the Increasing direction mostly but Japan is showing a decreasing trend over the years this might be due to technogical advancements or better management

### Were there any interesting or surprising interactions between features?
Green house gas emmision is a global phenomenon and the presence of green house gasses produced by industries are  increasing in many cases this was true as in cases of India and China but japan showed signs of decres too in certain sectors. 

#MODELING USING ARIMA MODEL
```{r echo = FALSE, message=FALSE, warning=FALSE , model_arima}
#Arima time series
sub1 <-subset(ghg,ghg$Country == 'India' )
subt <- subset(sub1,select = c( 'Year' ) )
sub <- ts(subset(sub1,select = c('TotalGHG'  ) ))


subt$cleanTghg <- tsclean(sub)

ggplot() +
  geom_line(data = subt, aes(x = Year, y = cleanTghg)) + 
  ylab('Cleaned TOTAL GHG')







```

 Calculating moving averages and ploting it

```{r echo=FALSE , message=FALSE, warning=FALSE}


subt$ghg_ma5 <- ma(subt$cleanTghg, order=5) #ma= moving average   5 years
subt$ghg_ma10 <- ma(subt$cleanTghg, order=10)# ma  decade
summary(subt$ghg_ma5)
summary(subt$ghg_ma10)
ggplot() +
  geom_line(data = subt, aes(x = Year, y =cleanTghg , colour = "Totalghg")) +
  geom_line(data = subt, aes(x = Year, y = subt$ghg_ma5,   
                             colour = "Moving Average for 5 years"))+
  geom_line(data = subt, aes(x = Year, y = subt$ghg_ma5, 
                             colour = " Moving Average for 10 years"))+
  ylab('GHG ')  




```

 Not much variation is seen here

 DECOMPOSITION

 TREND THE GHG emmision follows a increasing pattern overtime

 Seasonality yearly data no seasonality here  


```{r echo = TRUE}
tghg_ma <- ts(na.omit(subt$ghg_ma5), frequency=5) 
#taking 5 years as a frequeny and will build model around it

decomp = stl(tghg_ma, s.window="periodic" ,)
deseasonal_ghg <- seasadj(decomp)
plot(decomp)

#trend could easily be seen almost not too much noise 


#Arima model requires a stationary series

adf.test(tghg_ma, alternative = "stationary")

#Autocorrelation plots for chosing Arima parametres

Acf(tghg_ma, main='')

Pacf(tghg_ma, main='')

 


```


 Creating models to find best fit

```{r echo =TRUE}
m1 <- auto.arima(deseasonal_ghg, seasonal=FALSE)
m1
#model is ok because no major corelation can be seen from ACF andPACF checking for '6'
tsdisplay(residuals(m1), lag.max=10, main='(0,2,0) Model Residuals') 


m2 <- arima(deseasonal_ghg, order=c(1,1,6))

m2

tsdisplay(residuals(m2), lag.max=10, main='Seasonal Model Residuals')
#m2 model with  6 is better and shows no correlation

fcast <- forecast(m2, h=15)
plot(fcast)
```

  HOLDING VALUES AND INTRODUCING SEASONALITY

```{r echo=FALSE}

hold <- window(ts(deseasonal_ghg), start=2)
deseasonal_ghg
hold
fit_holdout <- arima(ts(deseasonal_ghg[-c(700:725)]), order=c(1,1,7))

fcast_holdout <- forecast(fit_holdout,h=15)

fcast_holdout
plot(x=fcast_holdout, main="Pridicting future values using forecast")


fit_seasonality = auto.arima(deseasonal_ghg, seasonal=TRUE)
fit_seasonality
fcast_s <- forecast(fit_seasonality, h=20)
plot(fcast_s)

```

### OPTIONAL: Did you create any models with your dataset? 
Discuss the \strengths and limitations of your model.

ARIMA (Autoregressive integrated moving average) is very widely used model time series analysis this model Also the Data was Non-seasonal in nature so non seasonal - Arima(p,d,q) model was used.
The model is built on a single variable to analyze its future values by using previous values .
The optimal model was found out to be ARIMA(1,1,6) [AR=1 , I=1 , MA=6] .

####STRENGTH-
Optimal for Univariate forecast of a Time series data
Uses lags(previous values) to help possible predict future values
80% confidence limits shaded in darker blue, and 95% in lighter blue.


####WEAKNESS-
The forecast broadens and becomes less relaible as time increase

 
------

# Final Plots and Summary


### Plot One
```{r echo=FALSE, Plot_One}



ggplot(aes(y= TotalGHG , x= year_bucket) ,
       data = ghga ) +
  geom_boxplot(aes(color = Country) )


```

### Description One

This plot groups data for some of the countries and compares thier Total GHG emmision this plot shows the huge disparitries of emmision b/w China and others. Japan is not showing any significant increase in emmision and its gg emmision in comparison to China and india is almost negligible india on the other hand is observing a stedy increase in emmision over the years.  

### Plot Two
```{r echo=FALSE, Plot_Two}


grid.arrange(arrangeGrob(pl1,pl2,pl3,pl4,pl5,pl6 , 
                         top = "Sectors VS Year comparison for some Countries", 
                         layout_matrix = lay) , ncol=1  )
  
```

### Description Two
The plot dipicts the trend followed by  some of the countries over the years in diffrent fields such as Transport , Manufacturing etc .China is mostly at the to in Emmision . The trend is in the Increasing direction mostly but Japan is showing a decreasing trend in variables such as Transport emmision , Emmision due to waste which is very rare given this dataset.


### Plot Three
```{r echo=FALSE, Plot_Three}
plot(x=fcast_holdout, main="Pridicting future values using forecast" , xlab = 'year[1 corresponds to 1990 and so on]' , ylab = 'TotalGHG Emmision')
 


```

### Description Three
Forecast of timeseries data of India pridicting  next 15 values for total ghg emissions. 
80% confidence limits shaded in darker blue, and 95% in lighter blue. Longer term forecasts will usually have more uncertainty, as the model gives further future values it will become more and more uncertain this could be seen as the  spread increases along the xaxis(time)

------

# Reflection

>Green House Gases are the major factors in promoting Global warming and is a global threat .Mostly Univariate Analysis was only useful for finding the nature and distribution of data not much Could be drawn out from it . Bivariate analysis helped in identifieng the anomalities / diffrent patterns in the Data This was very Useful as the data was highly corelated and it would mostly be a increasing linear Trend. Finnaly The Multivariate Analysis brought about some meaningful analytics and some infrences Could be drawn from it.
Time series Data was bit diffrent Than a Normal Data and hence modeling it was a chalange at first but with the right model it was very effect to Analayze and produce prictions for this Data 
The huge diffrences in emmision of countries like Japan And China Were Note worthy Though they also have diffrence in Population and Land-Mass But Japan had a significantly more GDP Percapita Than China.  
The Analysis of of Emmisions with Various Economic and social parametres could be done to produce more insights . 



